How Slow do my Transient Type Checks Go?

### 2019-08-17 paper.tex
Fixed conclusion

For each benchmark we highlighted one or two type annotations that appear to show a pattern for the typing lattice performance scatter plots, or in the case of SpectralNorm and Json had higher than usual columns. The identified types are those represented by the red and blue columns. The scatter plots are colour-coded accordingly: a red or blue circle represents configurations with the given type annotation present, but not both, purple circles represent configurations with both type annotations present, and grey circles represent those with neither. Though we exhaustively inspected such colour-coded scatter plot for all 1,775 type annotations across all 21 benchmarks, the only patterns we noticed are those shown in \ref{f:pattern}.

@@

In this paper we have investigated how benchmarks can be
repurposed to determine precisely which transient type checks are
resistant to the optimisations provided by a just-in-time compiling
virtual machine.  We observed that many of our benchmark results
conducted under the Takikawa protocol 
seemed to have a characteristic bimodal performance profile: some
of the benchmark configurations ran significantly slower than
the remaining configurations. We also observed trimodal profiles, as well as performance increases when type annotations where added.
By inspecting benchmark graphs colour coded for each individual type annotation in a program's source code, we were often able to notice patterns which suggested
that just one or two of the type annotations were likely responsible for
the bimodal performance.
However, this approach requires inspecting one graph per type annotation.

We also investigated another approach to help identify such annotations, for an individual benchmark, we can produce a single graph of the performances of individual configurations for each type annotation. 
In most cases both approaches identified the same type annotations as likely causing a significant effect on performance. In particular these type annotations seemed to be responsible for most of the overhead caused by Moth's transient type checking.

>>>>

For each benchmark we highlighted one or two type annotations that appear to show a pattern for the typing lattice performance scatter plots, or in the case of SpectralNorm and Json had higher than usual columns. The identified types are those represented by the red and blue columns. The scatter plots are colour-coded accordingly: a red or blue circle represents configurations with the given type annotation present, but not both, purple circles represent configurations with both type annotations present, and grey circles represent those with neither. Though we exhaustively inspected such colour coded scatter plots for all 1,775 type annotations across all 21 benchmarks, the only patterns we noticed are those shown in \ref{f:pattern}.

@@

In this paper we have investigated how benchmarks can be
repurposed to determine precisely which transient type checks are
resistant to the optimisations provided by a just-in-time compiling
virtual machine.  We observed that many of our benchmark results
conducted under the Takikawa protocol 
seemed to have a characteristic bimodal performance profile: some
of the benchmark configurations ran significantly slower than
the remaining configurations. We also observed trimodal profiles, as well as performance increases when type annotations where added.

By inspecting graphs of the performances of individual configurations for each type annotation, we can see type annotations that likely have a significant effect on performance. By then inspecting the typing lattice graphs colour coded for such type annotations, we were often able to notice patterns which suggested that just one or two of the type annotations were likely responsible for the bimodal performance. Though every type annotation that appeared to have a significant effect across the typing lattice also showed a singificant performance effect when no other typing annotations where present, the converse did not hold. In particular, the type annotations we identified appear to be responsible for most of the overhead caused by Moth's transient type checking.
